---
title: "Fidelity - Figures"
author: "Emily Shertzer"
date: "`r Sys.Date()`"
output: html_document
---

This script creates figures for the manuscript "Site fidelity of sagebrush songbirds largely invariant with cues of previous nest success or habitat disturbance across spatial scales".

Model outputs from:
RELEASE_Fidelity_landscape.Rmd
RELEASE_Fidelity_territory.Rmd
RELEASE_Fidelity_nestsite.Rmd

Data from: Shertzer, E.R., and Chalfoun, A.D., 2026, Mark recapture and habitat data for adult sagebrush songbirds tracked in Pinedale, Wyoming during 2020 - 2024: U.S. Geological Survey data release, https://doi.org/10.5066/P1EOPE4V.

Geolocator data from MoveBank project Sagebrush songbirds - University of Wyoming - 33-1299; DOI pending data review and release through the MoveBank repository.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo = FALSE}

# clear environment
rm(list = ls())
library(geosphere)
library(GeoLight)
library(changepoint)
library(BAStag)
library(devtools)
library(SGAT)
library(forecast)
library(MASS)
library(mvtnorm)
library(FLightR)
library(raster)
library(fields)
library(parallel)
library(readxl)
library(DT)
library(ggmap)
library(crosstalk)
library(htmltools)
library(leaflet)
library(grid)
library(assignR)
library(terra)
library(gridExtra)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggplot2)
library(sf)
library(ggspatial)
library(ggnewscale)
library(dplyr)
library(lubridate)
library(patchwork)
library(tidyverse)
# ggmap::register_google()  # Must use your own key
# register_stadiamaps(, write = FALSE) # Must use your own key

```

## Figure 2

The following code plots the breeding season locations of the three birds with geolocators that showed temporary emigration during one breeding season (two brsp, one sath).

  **Geo:** CB920<br>
  **Band number:** 291075323<br>
  **Start date:** 6/01/2021<br>
  **Recap date:** 5/29/2023<br>
  **Last/first resights:** 7/12/21 (No resights in 2022) <br>
  *This bird was tracked for 2 years with the same geolocator*

```{r CB920, echo=FALSE}
load("./FILEPATH/CB920RESULT_FlightR.Rdata")
Summary <- readRDS("./FILEPATH/CB920_stationary_prob0.05_min5.rds")

start <- Result$Results$Movement.results$Real.time[700]
stop <- Result$Results$Movement.results$Real.time[800]

dates <- data.frame(start = start,
                     stop = stop)

breed_ud_cb920 <- plot_util_distr(Result, dates = dates, geom_polygon.options = list(linetype = "blank"), color.palette = colorRampPalette("#55D6BE"), zoom = 5)

# Request the map
center <- c(-109.85206, 42.749746)
s <- paste0(
  "style=element:geometry%7Ccolor:0xf5f5f5",
  "&style=element:labels%7Cvisibility:off",
  "&style=element:labels.icon%7Cvisibility:off",
  "&style=element:labels.text.fill%7Ccolor:0x616161",
  "&style=element:labels.text.stroke%7Ccolor:0xf5f5f5",
  "&style=feature:administrative%7Celement:geometry%7Cvisibility:off",
  "&style=feature:administrative.country%7Celement:geometry.stroke%7Ccolor:0x000000%7Cvisibility:on",
  "&style=feature:administrative.province%7Celement:geometry.stroke%7Ccolor:0x8f8f8f%7Cvisibility:on",
  "&style=feature:poi%7Cvisibility:off",
  "&style=feature:road%7Cvisibility:off",
  "&style=feature:transit%7Cvisibility:off",
  "&style=feature:water%7Celement:geometry%7Ccolor:0xc9c9c9"
)


basemap <- get_googlemap(
  center = center,
  zoom = 5,
  scale = 2,
  style = s,
)

sf <- breed_ud_cb920$res_buffers %>%
  dplyr::mutate(geometry = sf::st_sfc(b)) %>%   
  dplyr::select(-b) %>%                         
  st_as_sf(crs = 4326)  

sf_polys <- sf %>% st_cast("POLYGON")

poly_df <- sf_polys %>%
  st_coordinates() %>%
  as.data.frame() %>%
  rename(lon = X, lat = Y) %>%
  mutate(group = interaction(L1, L2)) 

cb920 <- ggmap(basemap) +
  geom_polygon(data = poly_df,
               aes(x = lon, y = lat, group = group),
               fill = "#55D6BE", alpha = 0.4) +
  scale_x_continuous(limits = c(-115, -105), expand = c(0,0)) +
  scale_y_continuous(limits = c(39, 48.5), expand = c(0,0)) +
  theme_minimal() +
  theme(axis.text = element_text(),
        legend.position = "none")
```

```{r CB924, echo=FALSE}

load("./FILEPATH/CB924RESULT_FlightR.Rdata")
Summary <- readRDS("./FILEPATH/CB924_stationary_prob0.2_min5.rds")

#breeding location in the second year
start <- Result$Results$Movement.results$Real.time[700]  
#basing this on the Summary$Potential_stat_periods
stop <- Result$Results$Movement.results$Real.time[800]

dates <- data.frame(start = start,
                     stop = stop)

breed_ud_cb924 <- plot_util_distr(Result, dates = dates, geom_polygon.options = list(linetype = "blank"), color.palette = colorRampPalette("#55D6BE"), zoom = 6)

sf <- breed_ud_cb924$res_buffers %>%
  dplyr::mutate(geometry = sf::st_sfc(b)) %>%   
  dplyr::select(-b) %>%                         
  st_as_sf(crs = 4326)  

sf_polys <- sf %>% st_cast("POLYGON")

poly_df <- sf_polys %>%
  st_coordinates() %>%
  as.data.frame() %>%
  rename(lon = X, lat = Y) %>%
  mutate(group = interaction(L1, L2))  # L1 = polygon ID, L2 = ring ID

cb924 <- ggmap(basemap) +
  geom_polygon(data = poly_df,
               aes(x = lon, y = lat, group = group),
               fill = "#55D6BE", alpha = 0.4) +
  #geom_text(x = -109.7785, y = 42.3, label = "*", size = 6, color = "black") +
  scale_x_continuous(limits = c(-115, -105), expand = c(0,0)) +
  scale_y_continuous(limits = c(39, 48.5), expand = c(0,0)) +
  theme_minimal() +
  theme(axis.text = element_text(),
        legend.position = "none")


```


```{r CB885, echo=FALSE}
load("./FILEPATH/CB885RESULT_FlightR_2calibperiods.Rdata")
Summary <- readRDS("./FILEPATH/CB885_stationary_2calib_prob0.1_min3.rds")

distance <- (distHaversine(cbind(-109.7785, 42.3), cbind(Summary$Stationary.periods$Medianlon[24], Summary$Stationary.periods$Medianlat[24]))/1000) 

start <- Result$Results$Movement.results$Real.time[650]  #basing this on the Summary$Potential_stat_periods - period following breeding
stop <- Result$Results$Movement.results$Real.time[780]

dates <- data.frame(start = start,
                     stop = stop)

breed_ud_cb885 <- plot_util_distr(Result, dates = dates, geom_polygon.options = list(linetype = "blank"), color.palette = colorRampPalette("#DDA448"), zoom = 6)


sf <- breed_ud_cb885$res_buffers %>%
  dplyr::mutate(geometry = sf::st_sfc(b)) %>%   
  dplyr::select(-b) %>%                         
  st_as_sf(crs = 4326)  

sf_polys <- sf %>% st_cast("POLYGON")

poly_df <- sf_polys %>%
  st_coordinates() %>%
  as.data.frame() %>%
  rename(lon = X, lat = Y) %>%
  mutate(group = interaction(L1, L2))  # L1 = polygon ID, L2 = ring ID

cb885 <- ggmap(basemap) +
  geom_polygon(data = poly_df,
               aes(x = lon, y = lat, group = group),
               fill = "#DDA448", alpha = 0.4) +
  #geom_text(x = -109.7785, y = 42.3, label = "*", size = 6, color = "black") +
  scale_x_continuous(limits = c(-115, -105), expand = c(0,0)) +
  scale_y_continuous(limits = c(39, 48.5), expand = c(0,0)) +
  theme_minimal() +
  theme(axis.text = element_text(),
        legend.position = "none")


```

## Figure 3
Figure to compare model results across landscape, territory, and nest site scales.


```{r betafig, echo=FALSE}
wd <- c("")  #working directory

#age models....
load(file=paste0(wd, "FILEPATH")) #landscape mod
load(file=paste0(wd, "FILEPATH")) #territory mod
load(file=paste0(wd, "FILEPATH")) #nest site mod

species_cols <- c(
  "BRSP" = "#55D6BE",
  "SABS" = "#FA003F",
  "SATH" = "#DDA448"
)


#landscape first

mcmc <- landscape$summary

beta <- mcmc %>%
  as.data.frame() %>%
  mutate(param = rownames(.)) %>%
  filter(str_detect(param, "^beta[123456]_gPP")) %>%
  # extract covariate number + species ID
  mutate(
    cov = str_extract(param, "beta\\d"),
    spp = str_extract(param, "\\[(\\d)\\]$") %>% str_remove_all("\\[|\\]") %>% as.numeric()
  )

beta <- beta %>%
  mutate(cov_label = dplyr::recode(cov,
                                   "beta1" = "Sex",
                                   "beta2" = "Disturbance",
                                   "beta3" = "Shrub Height",
                                   "beta4" = "Nest Fail Rate",
                                   "beta5" = "Shrub Cover", 
                                   "beta6" = "Age Class"),
         spp_label = dplyr::recode(as.character(spp),
                                   "1" = "BRSP",
                                   "2" = "SABS",
                                   "3" = "SATH"),
         sig = if_else(`2.5%` > 0 | `97.5%` < 0, "Non-zero", "Overlaps 0")
  ) %>%
  mutate(spp_label = factor(spp_label, levels = c("BRSP", "SABS", "SATH")))

beta_plot <- beta %>%
  mutate(cov_label = factor(cov_label, levels = c("Shrub Cover", "Shrub Height", 
                                              "Nest Fail Rate", "Disturbance", 
                                              "Sex", "Age Class")),
    draw = !(cov_label == "Age Class" & spp_label == "SATH"),
    mean_plot = if_else(draw, mean, NA_real_),
    lwr_plot  = if_else(draw, `2.5%`, NA_real_),
    upr_plot  = if_else(draw, `97.5%`, NA_real_),
    label     = if_else(draw & f > 0.9, "*", NA_character_)
  )

landscape <- ggplot(beta_plot, aes(x = spp_label, y = mean_plot, 
                              ymin = lwr_plot, ymax = upr_plot,
                              color = spp_label)) +
  geom_pointrange(size = 0.8, position = position_dodge(width = 0.5)) +
  geom_text(aes(label = label, y = lwr_plot), 
          vjust = 1.5, size = 6, show.legend = FALSE) +
  facet_wrap(~ cov_label, scales = "free_x", nrow = 1) +
  scale_color_manual(values = species_cols) +
  geom_hline(yintercept = 0, linetype = 2, linewidth = 0.7) +
  labs(x = "Species", y = "Effect on Site Fidelity (β)", color = "") +   #Effect on  (β)
  theme_bw(base_size = 14) +
  theme(
    strip.text = element_text(size = 16),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16),
    legend.position = "none"
  )

# Territory Scale
mcmc <- territory$summary

beta <- mcmc %>%
  as.data.frame() %>%
  mutate(param = rownames(.)) %>%
  filter(str_detect(param, "^beta[123456]")) %>%
  # extract covariate number + species ID
  mutate(
    cov = str_extract(param, "beta\\d"),
    spp = str_extract(param, "\\[(\\d)\\]$") %>% str_remove_all("\\[|\\]") %>% as.numeric()
  )

#flipping this one, so that betas are describing "site fidelity" rather than movement
beta$mean <- -beta$mean
beta$`2.5%` <- -(beta$`2.5%`)
beta$`97.5%` <- -beta$`97.5%`

beta <- beta %>%
  mutate(cov_label = dplyr::recode(cov,
                                   "beta1" = "Sex",
                                   "beta2" = "Nest Fail Rate",
                                   "beta3" = "Disturbance",
                                   "beta4" = "Shrub Height",
                                   "beta5" = "Shrub Cover",
                                   "beta6" = "Age Class"),
         spp_label = dplyr::recode(as.character(spp),
                                   "1" = "BRSP",
                                   "2" = "SATH",
                                   "3" = "SABS"),
         sig = if_else(`2.5%` < 0 | `97.5%` > 0, "Non-zero", "Overlaps 0") #flipped
  ) %>%
  mutate(spp_label = factor(spp_label, levels = c("BRSP", "SABS", "SATH")))

beta_plot <- beta %>%
  mutate(cov_label = factor(cov_label, levels = c("Shrub Cover", "Shrub Height", 
                                              "Nest Fail Rate", "Disturbance", 
                                              "Sex", "Age Class")),
    draw = !(cov_label == "Age Class" & spp_label == "SATH"),
    mean_plot = if_else(draw, mean, NA_real_),
    lwr_plot  = if_else(draw, `2.5%`, NA_real_),
    upr_plot  = if_else(draw, `97.5%`, NA_real_),
    label     = if_else(draw & f > 0.9, "*", NA_character_)
  )

local <- ggplot(beta_plot, aes(x = spp_label, y = mean_plot, 
                              ymin = lwr_plot, ymax = upr_plot,
                              color = spp_label)) +
  geom_pointrange(size = 0.8, position = position_dodge(width = 0.5)) +
  geom_text(aes(label = label, y = upr_plot), 
          vjust = 1.5, size = 6, show.legend = FALSE) +
  facet_wrap(~ cov_label, scales = "free_x", nrow = 1) +
  scale_color_manual(values = species_cols) +
  geom_hline(yintercept = 0, linetype = 2, linewidth = 0.7) +
  labs(x = "Species", y = "Effect on Site Fidelity (β)", color = "") +   #Effect on  (β)
  theme_bw(base_size = 14) +
  theme(
    strip.text = element_text(size = 16),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16),
    legend.position = "none"
  )

# adding in nest site scale

mcmc <- nestsite$summary

beta <- mcmc %>%
  as.data.frame() %>%
  mutate(param = rownames(.)) %>%
  filter(str_detect(param, "^beta[123456]")) %>%
  # extract covariate number + species ID
  mutate(
    cov = str_extract(param, "beta\\d"),
    spp = str_extract(param, "\\[(\\d)\\]$") %>% str_remove_all("\\[|\\]") %>% as.numeric()
  )

#flipping this one, so that betas are describing "site fidelity" rather than movement
beta$mean <- -beta$mean
beta$`2.5%` <- -(beta$`2.5%`)
beta$`97.5%` <- -beta$`97.5%`

beta <- beta %>%
  mutate(cov_label = dplyr::recode(cov,
                                   "beta1" = "Sex",
                                   "beta2" = "Nest Fate",
                                   "beta3" = "Disturbance",
                                   "beta4" = "Shrub Height",
                                   "beta5" = "Shrub Cover",
                                   "beta6" = "Age Class"),
         spp_label = dplyr::recode(as.character(spp),
                                   "1" = "BRSP",
                                   "2" = "SATH",
                                   "3" = "SABS"),
         sig = if_else(`2.5%` > 0 | `97.5%` < 0, "Non-zero", "Overlaps 0")
  ) %>%
  mutate(spp_label = factor(spp_label, levels = c("BRSP", "SABS", "SATH")))

beta_plot <- beta %>%
  mutate(cov_label = factor(cov_label, levels = c("Shrub Cover", "Shrub Height", 
                                              "Nest Fate", "Disturbance", 
                                              "Sex", "Age Class")),
    draw = !(cov_label == "Age Class" & spp_label == "SATH"),
    mean_plot = if_else(draw, mean, NA_real_),
    lwr_plot  = if_else(draw, `2.5%`, NA_real_),
    upr_plot  = if_else(draw, `97.5%`, NA_real_),
    label     = if_else(draw & f > 0.9, "*", NA_character_)
  )

nest <- ggplot(beta_plot, aes(x = spp_label, y = mean_plot, 
                              ymin = lwr_plot, ymax = upr_plot,
                              color = spp_label)) +
  geom_pointrange(size = 0.8, position = position_dodge(width = 0.5)) +
  geom_text(aes(label = label, y = upr_plot), 
          vjust = 1.5, size = 6, show.legend = FALSE) +
  coord_cartesian(ylim = c(min(na.omit(beta_plot$upr_plot)) - 0.6, max(na.omit(beta_plot$lwr_plot)))) +
  facet_wrap(~ cov_label, scales = "free_x", nrow = 1) +
  scale_color_manual(values = species_cols) +
  geom_hline(yintercept = 0, linetype = 2, linewidth = 0.7) +
  labs(x = "Species", y = "Effect on Site Fidelity (β)", color = "") +   #Effect on  (β)
  theme_bw(base_size = 14) +
  theme(
    strip.text = element_text(size = 16),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 16),
    legend.position = "bottom",       
    legend.direction = "horizontal",
    legend.box = "horizontal",     
    legend.text = element_text(size = 16),
  )

landscape <- landscape + theme(plot.title = element_text(size = 16))
local     <- local     + theme(plot.title = element_text(size = 16))
nest      <- nest      + theme(plot.title = element_text(size = 16))

final_plot <- 
  landscape + 
  labs(title = "Landscape") +
  local + 
  labs(title = "Territory") +
  nest + 
  labs(title = "Nest-Site") +
  plot_layout(ncol = 1) +
  plot_annotation(theme = theme(plot.title = element_text(size = 16)))

```

## Figure 4

Histograms to visualize typically territory shifts for each species, and movement toward/away from disturbance for individuals tracked over multiple years.

```{r fig4, echo=FALSE}

species_cols <- c(
  "BRSP" = "#55D6BE",
  "SABS" = "#FA003F",
  "SATH" = "#DDA448"
)

#bring in & format data for this fig

surv <- read.csv("./FILEPATH/SageSongbirdSurv_Final.csv", colClasses = c(ch = "character"))

test <- surv %>%
  dplyr::select(BandNumber, Species, Sex, plot,
                mean_utme_2020, mean_utmn_2020,
                mean_utme_2021, mean_utmn_2021, 
                mean_utme_2022, mean_utmn_2022,
                mean_utme_2023, mean_utmn_2023, 
                mean_utme_2024, mean_utmn_2024, 
                kmdistur2020, kmdistur2021, kmdistur2022, kmdistur2023, kmdistur2024) %>%
  dplyr::mutate(
    dist2020_2021 = as.numeric(sqrt((mean_utme_2021 - mean_utme_2020)^2 +
                                      (mean_utmn_2021 - mean_utmn_2020)^2)),
    dist2021_2022 = as.numeric(sqrt((mean_utme_2022 - mean_utme_2021)^2 +
                                      (mean_utmn_2022 - mean_utmn_2021)^2)),
    dist2022_2023 = as.numeric(sqrt((mean_utme_2023 - mean_utme_2022)^2 +
                                      (mean_utmn_2023 - mean_utmn_2022)^2)),
    dist2023_2024 = as.numeric(sqrt((mean_utme_2024 - mean_utme_2023)^2 +
                                      (mean_utmn_2024 - mean_utmn_2023)^2))
  )

plotdat <- test %>%
  transmute(
    BandNumber,
    Species,
    Sex,
    plot,
    year = 2020, 
    kmdistur = kmdistur2020,
    distance = dist2020_2021) %>%
  bind_rows(
    test %>%
      transmute(
        BandNumber,
        Species,
        Sex,
        plot,
        year = 2021,
        kmdistur = kmdistur2021,
        distance = dist2021_2022),
    test %>%
      transmute(
        BandNumber,
        Species,
        Sex,
        plot,
        year = 2022,
        kmdistur = kmdistur2022,
        distance = dist2022_2023),
    test %>%
      transmute(
        BandNumber,
        Species,
        Sex,
        plot,
        year = 2023,
        kmdistur = kmdistur2023,
        distance = dist2023_2024))

#prep data for jags

mod_dat <- plotdat %>%
  dplyr::filter(!is.na(distance)) %>%
  dplyr::mutate(sex = case_when(Sex == "M" ~ 1, Sex == "F" ~ 0, Sex == "U" ~ NA_real_))

mod_dat <- mod_dat %>%
  mutate(
    species_id = as.integer(as.factor(Species)),
    band_id = as.integer(as.factor(BandNumber)),
    year_id = as.integer(as.factor(year)))
x_max <- max(mod_dat$distance, na.rm = TRUE)

hist1 <- ggplot(mod_dat[mod_dat$Species == "BRSP", ], aes(x = distance)) +
  annotate(
    "rect",
    xmin = 131.55 * 2, xmax = 165.09 * 2,
    ymin = -Inf, ymax = Inf,
    fill = "grey70", alpha = 0.4
  ) +
    geom_vline(
    xintercept = (150.59*2),  # median territory size (diameter)
    linetype = "dotted",
    linewidth = 1,
    color = "black"
  ) +
  geom_histogram(
    bins = 30,
    fill = "#55D6BE",      
    color = "#55D6BE",       
    alpha = 0.80
  ) +
  coord_cartesian(xlim = c(0, x_max)) +
  labs(x = NULL, y = "Count") +
  theme_classic(base_size = 16) +
  theme(
    axis.text.x  = element_blank(),
    axis.title.y = element_blank())


hist2 <- ggplot(mod_dat[mod_dat$Species == "SABS", ], aes(x = distance)) +
    annotate(
    "rect",
    xmin = 141.05 * 2, xmax = 179.52 * 2,
    ymin = -Inf, ymax = Inf,
    fill = "grey70", alpha = 0.4
  ) +
    geom_vline(
    xintercept = (162.05*2),  # median territory size (diameter)
    linetype = "dotted",
    linewidth = 1,
    color = "black"
  ) +
  geom_histogram(
    bins = 30,
    fill = "#FA003F",     
    color = "#FA003F",      
    alpha = 0.80
  ) +
  coord_cartesian(xlim = c(0, x_max)) +
  labs(x = NULL, y = "Count") +
  theme_classic(base_size = 16) +
  theme(axis.text.x  = element_blank())


hist3 <- ggplot(mod_dat[mod_dat$Species == "SATH", ], aes(x = distance)) +
    annotate(
    "rect",
    xmin = 151.91 * 2, xmax = 181.18 * 2,
    ymin = -Inf, ymax = Inf,
    fill = "grey70", alpha = 0.4
  ) +
    geom_vline(
    xintercept = (166.89*2),  # median territory size (diameter)
    linetype = "dotted",
    linewidth = 1,
    color = "black"
  ) +
  geom_histogram(
    bins = 30,
    fill = "#DDA448",     
    color = "#DDA448",      
    alpha = 0.70
  ) +
  coord_cartesian(xlim = c(0, x_max)) +
  labs(
    x = "Distance territory shifted (m)",
    y = "Count"
  ) +
  theme_classic(base_size = 16) +
  theme(axis.title.x = element_text(face = "plain"),
        axis.title.y = element_blank())

left_col <- hist1 / hist2 / hist3

plotdat2 <- plotdat %>%
  dplyr::filter(!is.na(kmdistur)) %>%
  dplyr::mutate(Species = factor(Species, levels = c("BRSP", "SABS", "SATH"))) %>%
  group_by(BandNumber) %>%
  filter(n() >= 2) %>%   # keep only individuals with ≥ 2 years
  dplyr::mutate(year_since_first = year - min(year),
                dist_rel = kmdistur - first(kmdistur) ) %>%
  ungroup() 

plot1 <- ggplot(plotdat2, aes(x = year_since_first, y = dist_rel,
                     group = BandNumber, color = Species)) +
  geom_line(size = 1, alpha = 0.7) +
  geom_point(alpha = 0.6) +
  labs(x = "Years since first capture",
       y = expression("Relative "*Delta*" disturbance (proportion)")) +
  scale_color_manual(values = c(
    "BRSP" = "#55D6BE",
    "SABS" = "#FA003F",
    "SATH" = "#DDA448"
  )) +
  theme_classic(base_size = 16)

combined_plot <- left_col | plot1

```

## Table 1

Post-hoc pairwise credible differences between years in site fidelity of Brewer’s sparrows, sagebrush sparrows, and sage thrashers.

```{r time, echo=FALSE}

#landscape fidelity model...
beta0_samps <- landscape$sims.list$beta0_gPP

#territory fidelity model... (Sags and sath are switched)
beta0_samps <- territory$sims.list$beta0

#nest fidelity model... (SAGS and SATH are switched)
beta0_samps <- nestsite$sims.list$beta0

yrs <- 1:4
pairs <- combn(yrs, 2, simplify = FALSE)
#brsp
results <- lapply(pairs, function(p) {
  d <- beta0_samps[, p[1], 1] - beta0_samps[, p[2], 1]
  data.frame(
    contrast = paste0("beta0[", p[1], ",1] - beta0[", p[2], ",1]"),
    median = median(d),
    lwr = quantile(d, 0.025),
    upr = quantile(d, 0.975),
    prob_gt0 = bayestestR::pd(d)
  )
})

do.call(rbind, results)

#sags
results <- lapply(pairs, function(p) {
  d <- beta0_samps[, p[1], 2] - beta0_samps[, p[2], 2]
  data.frame(
    contrast = paste0("beta0[", p[1], ",2] - beta0[", p[2], ",2]"),
    median = median(d),
    lwr = quantile(d, 0.025),
    upr = quantile(d, 0.975),
    prob_gt0 = bayestestR::pd(d)
  )
})

do.call(rbind, results)

#sath
results <- lapply(pairs, function(p) {
  d <- beta0_samps[, p[1], 3] - beta0_samps[, p[2], 3]
  data.frame(
    contrast = paste0("beta0[", p[1], ",3] - beta0[", p[2], ",3]"),
    median = median(d),
    lwr = quantile(d, 0.025),
    upr = quantile(d, 0.975),
    prob_gt0 = bayestestR::pd(d)
  )
})

do.call(rbind, results)

```
