---
title: "Fidelity_landscape"
author: "Emily Shertzer"
date: "`r Sys.Date()`"
output: html_document
---

## Overview 

This script creates a multi-state robust design mark recapture model, used to estimated landscape-scale fidelity for the manuscript "Site fidelity of sagebrush songbirds largely invariant with cues of previous nest success or habitat disturbance across spatial scales".

Covariates: disturbance, nest failure rates (plot/year), shrub height, shrub cover, tag status (mark), crew (observer)

Data: Shertzer, E.R., and Chalfoun, A.D., 2026, Mark recapture and habitat data for adult sagebrush songbirds tracked in Pinedale, Wyoming during 2020 - 2024: U.S. Geological Survey data release, https://doi.org/10.5066/P1EOPE4V.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo = FALSE}
# clear environment
rm(list = ls())

require(jagsUI)
library(MCMCpack)
library(abind)
library(tidyverse)
library(dplyr)
library(AHMbook)
library(lubridate)
library(ggplot2)
library(loo)
#no sci notation
options(scipen = 999)
```

## Data prep

```{r data_prep}
surv <- read.csv("./SageSongbirdSurv_Final.csv", colClasses = c(ch = "character"))

# adding two dummy secondary periods (2020 only had 1 real secondary period)
surv$ch <- paste("00", surv$ch, sep = "")

#now, removing birds that were captured in the last primary period
surv <- surv %>%
  dplyr::filter(ch != "000000000000001") %>%
  dplyr::filter(!ch %in% c("000000000000111", "000000000000110", "000000000000101", "000000000000011", "000000000000100", "000000000000010"))

#Data prep matrices for the model
species <- surv$Species
species[species == "BRSP"] <- 1
species[species == "SABS"] <- 2
species[species == "SATH"] <- 3
species <- as.numeric(species)
n.ind.s <- table(species)

#now, making a matrix of the ch for the model
ch <- apply((do.call("rbind",strsplit(surv$ch,""))),2,as.numeric)

#creating list of primary periods and f_ann (primary period of first capture)
ch_ann<-cbind(ch[,3],1*rowSums(ch[,4:6])>0,1*rowSums(ch[,7:9])>0, 1*rowSums(ch[,10:12])>0, 1*rowSums(ch[,13:15])>0)
f_ann<-apply(ch_ann,1,which.max)
#convert ch_ann to 1 and 2s; 1= seen, 2 =not seen
ch_ann[ch_ann == 0] <- 2

# Observation matrix

n.ind <- nrow(ch)      # number of individuals
n.prim <- 5            # number of primary periods
n.sec <- rep(3, n.prim)
n.sec[1] <- 1          #first primary only had 1 secondary period
n.spp <- 3

obs <- array(NA, dim = c(n.ind, n.prim, max(n.sec)))

# Fill the obs array
for (i in 1:n.ind) {
  for (t in 1:n.prim) {
    cols <- ((t - 1) * n.sec[t] + 1):(t * n.sec[t])  
    obs[i, t, 1:n.sec[t]] <- ch[i, cols]
  }
}

obs[, 1, 1:2] <- NA   # mark secondary 1 and 2 in primary 1 as missing

#adding in the species dimension
ch_array  <- array(NA, dim = c(max(n.ind.s), n.prim, n.spp))
obs_array <- array(NA, dim = c(max(n.ind.s), n.prim, 3, n.spp))


for (s in 1:n.spp) {
  ind_s <- which(species == s)     # which rows belong to species s
  n_s   <- length(ind_s)           # number of individuals of species s
  
  ch_array[1:n_s, , s] <- as.matrix(ch_ann[ind_s, ])
  obs_array[1:n_s, , ,s] <- as.matrix(obs[ind_s,,])
}


#sex: M = 1, F = 0, U = NA
sex <- replace(1*(surv$Sex=="M"),surv$Sex=="U",NA)

#prepping crew covariate (for detection)
crew<- cbind(surv$site2020, surv$site2021, surv$site2022, surv$site2023, surv$site2024)
all_crew <- sort(unique(as.vector(crew)))
crew_map <- setNames(seq_along(all_crew), all_crew)
crew <- matrix(crew_map[as.character(crew)], nrow = nrow(crew))
n.crew <- length(all_crew)

#prepping disturbance covariate
disturb<- cbind(surv$kmdistur2020, surv$kmdistur2021, surv$kmdistur2022, surv$kmdistur2023, surv$kmdistur2024)
disturb_scale<- scale(disturb, center=TRUE, scale=TRUE) #scale it

#prepping frate covariate (plot/year level nest failure rate)
frate <- cbind(surv$frate2020, surv$frate2021, surv$frate2022, surv$frate2023, surv$frate2024)
frate_scale<- scale(frate, center=TRUE, scale=TRUE) #scale it

#prepping veg covariates 
vegheight <- cbind(surv$shrub_height_mean2020, surv$shrub_height_mean2021, surv$shrub_height_mean2022, surv$shrub_height_mean2023, surv$shrub_height_mean2024)
vegheight_scale<- scale(vegheight, center=TRUE, scale=TRUE) #scale it

vegcov <- cbind(surv$veg_cover_mean2020, surv$veg_cover_mean2021, surv$veg_cover_mean2022, surv$veg_cover_mean2023, surv$veg_cover_mean2024)
vegcov_scale<- scale(vegcov, center=TRUE, scale=TRUE) #scale it

#fix ages.... so that AHY is NA in first year
surv <- surv %>%
  dplyr::mutate(age2020 = dplyr::if_else(Age == 1 & Year == 2020, NA, age2020),
                age2021 = dplyr::if_else(Age == 1 & Year == 2021, NA, age2021),
                age2022 = dplyr::if_else(Age == 1 & Year == 2022, NA, age2022),
                age2023 = dplyr::if_else(Age == 1 & Year == 2023, NA, age2023),
                age2024 = dplyr::if_else(Age == 1 & Year == 2024, NA, age2024))

age <- cbind(surv$age2020, surv$age2021, surv$age2022, surv$age2023, surv$age2024)
#bin age into 0's and 1s (0 = SY, 1 = ASY)
age <- ifelse(is.na(age), NA,
              ifelse(age == 2, 0, 1))

#prep tag data data
mark <- cbind(surv$marker2020, surv$marker2021, surv$marker2022, surv$marker2023, surv$marker2024)

mark[is.na(mark) | mark == ""] <- "N"
mark[mark == "N"] <- 1
mark[mark == "GEO"] <- 2
mark[mark == "LT"] <- 3
mark <- matrix(as.numeric(mark), nrow = nrow(mark))

# prep plot data
plot <- surv$plot
plot[plot == "J0"] <- 1
plot[plot == "J11"] <- 2
plot[plot == "J17"] <- 3
plot[plot == "P0"] <- 4
plot[plot == "P0 OFF"] <- 4
plot[plot == "P8"] <- 5
plot[plot == "P34"] <- 6
plot[plot == "P38"] <- 7
plot <- as.numeric(plot)

#prepping all of these covariates for multi-species dimensions....
first_array  <- array(NA, dim = c(max(n.ind.s), n.spp))
sex_array  <- array(NA, dim = c(max(n.ind.s), n.spp))
plot_array  <- array(NA, dim = c(max(n.ind.s), n.spp))
crew_array  <- array(NA, dim = c(max(n.ind.s), n.prim, n.spp))
disturb_array  <- array(NA, dim = c(max(n.ind.s), n.prim, n.spp))
age_array  <- array(NA, dim = c(max(n.ind.s), n.prim, n.spp))
mark_array  <- array(NA, dim = c(max(n.ind.s), n.prim, n.spp))
frate_array  <- array(NA, dim = c(max(n.ind.s), n.prim, n.spp))
vegheight_array  <- array(NA, dim = c(max(n.ind.s), n.prim, n.spp))
vegcov_array  <- array(NA, dim = c(max(n.ind.s), n.prim, n.spp))

for (s in 1:n.spp) {
  ind_s <- which(species == s)     # which rows belong to species s
  n_s   <- length(ind_s)           # number of individuals of species s
  
  first_array[1:n_s, s] <- as.matrix(f_ann[ind_s])
  sex_array[1:n_s, s] <- as.matrix(sex[ind_s])
  plot_array[1:n_s, s] <- as.matrix(plot[ind_s])
  crew_array[1:n_s, ,s] <- as.matrix(crew[ind_s,])
  disturb_array[1:n_s, ,s] <- as.matrix(disturb_scale[ind_s,])
  age_array[1:n_s, ,s] <- as.matrix(age[ind_s,])
  mark_array[1:n_s, ,s] <- as.matrix(mark[ind_s,])
  frate_array[1:n_s, ,s] <- as.matrix(frate_scale[ind_s,])
  vegheight_array[1:n_s, ,s] <- as.matrix(vegheight_scale[ind_s,])
  vegcov_array[1:n_s, ,s] <- as.matrix(vegcov_scale[ind_s,])
}

#b/c of small sample sizes and effects from those, making all SATH ages into NA:
age_array[,,3] <- NA 

```

## Collinearity check

Functions from:
Mixed effects models and extensions in ecology with R. (2009).
Zuur, AF, Ieno, EN, Walker, N, Saveliev, AA, and Smith, GM. Springer.

```{r collinearity, echo=FALSE}
panel.cor <- function(x, y, digits=1, prefix="", cex.cor = 6)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  r1=cor(x,y,use="pairwise.complete.obs")
  r <- abs(cor(x, y,use="pairwise.complete.obs"))
  txt <- format(c(r1, 0.123456789), digits=digits)[1]
  txt <- paste(prefix, txt, sep="")
  if(missing(cex.cor)) { cex <- 0.9/strwidth(txt) } else {
    cex = cex.cor}
  text(0.5, 0.5, txt, cex = cex * r)
}

cor_check <- surv %>%
  dplyr::select(BandNumber, kmdistur2021, kmdistur2022, kmdistur2023,
                frate2021, frate2022, frate2023,
                shrub_height_mean2021, shrub_height_mean2022, shrub_height_mean2023,
                veg_cover_mean2021, veg_cover_mean2022, veg_cover_mean2023,
                plot) %>%
  pivot_longer(
    cols = c(kmdistur2021, kmdistur2022, kmdistur2023,
             frate2021, frate2022, frate2023,
             shrub_height_mean2021, shrub_height_mean2022, shrub_height_mean2023,
             veg_cover_mean2021, veg_cover_mean2022, veg_cover_mean2023),
    names_to = c(".value","year"),
    names_pattern = "(.*)(\\d{4})"
  ) %>%
  mutate(plot = as.factor(plot))

MyVar <- c("kmdistur", "frate", "shrub_height_mean", "veg_cover_mean", "plot")

pairs(cor_check[,MyVar], lower.panel = panel.cor)

#Plot is a categorical variable, so collinearity was assessed visually for plot

```

## RD Model

```{r rd_model, echo=FALSE}
sink("spp_multiveg_age_markov.jags")
cat("
model {

########################################################################## 
#--------------------------------------
# Parameters:
# phi: survival probability
# p: detection probability
# gPP: temporary emigration parameter ( probability of remaining onsite, conditional on survival)
# gP: site immigration parameter (probability of returning onsite, conditional on survival)
#--------------------------------------
# States (S):
# 1 alive - observable (on site)
# 2 alive - nonobservable (off site)
# 3 dead
# Observations (O):
# 1 seen 
# 2 not seen
#--------------------------------------
##########################################################################


# priors - data imputation
psi ~ dunif(0,1)  #prior for sex ratio
psi2 ~ dunif(0,1)  #prior for age ratio

mu_disturb ~ dnorm(0, 0.01)
tau_disturb <- pow(sd_disturb, -2)
sd_disturb ~ dunif(0, 5)

mu_frate ~ dnorm(0, 0.01)   
tau_frate <- pow(sd_frate, -2)
sd_frate ~ dunif(0, 5)

mu_vegheight ~ dnorm(0, 0.01)    
tau_vegheight <- pow(sd_vegheight, -2)
sd_vegheight ~ dunif(0, 5)

mu_vegcov ~ dnorm(0, 0.01)
tau_vegcov <- pow(sd_vegcov, -2)
sd_vegcov ~ dunif(0, 5)

#phi priors

mu_beta1_phi ~ dnorm(0, 0.01)   # effect of sex
mu_beta2_phi ~ dnorm(0, 0.01)   # effect of disturbance
mu_beta3_phi ~ dnorm(0, 0.01)   # effect of veg height
mu_beta4_phi ~ dnorm(0, 0.01)   # effect of frate
mu_beta5_phi ~ dnorm(0, 0.01)   # effect of veg cover

sigma_beta1_phi ~ dunif(0, 5)
tau_beta1_phi <- pow(sigma_beta1_phi, -2)
sigma_beta2_phi ~ dunif(0, 5)
tau_beta2_phi <- pow(sigma_beta2_phi, -2)
sigma_beta3_phi ~ dunif(0, 5)
tau_beta3_phi <- pow(sigma_beta3_phi, -2)
sigma_beta4_phi ~ dunif(0, 5)
tau_beta4_phi <- pow(sigma_beta4_phi, -2)
sigma_beta5_phi ~ dunif(0, 5)
tau_beta5_phi <- pow(sigma_beta5_phi, -2)

for (s in 1:n.spp){
  beta1_phi[s] ~ dnorm(mu_beta1_phi, tau_beta1_phi) # prior for effect of sex on phi
  beta2_phi[s] ~ dnorm(mu_beta2_phi, tau_beta2_phi) #prior for effect of disturbance on phi
  beta3_phi[s] ~ dnorm(mu_beta3_phi, tau_beta3_phi)  # effect of veg height
  beta4_phi[s] ~ dnorm(mu_beta4_phi, tau_beta4_phi)  # effect of frate
  beta5_phi[s] ~ dnorm(mu_beta5_phi, tau_beta5_phi)  # effect of veg cover
}

#prior for the effect of marker (tag) on phi
alpha2_phi[1] <- 0 #reference level (No marker)
for (m in 2:3) {
  alpha2_phi[m] ~ dnorm(0, (1/2.25))   
}

#gamma priors: sex + disturb + frate
for(s in 1:n.spp){
beta1_gPP[s] ~ dnorm(mu_beta1_gPP, tau_beta1_gPP)   # prior for effect of sex on gammaPP
beta2_gPP[s] ~ dnorm(mu_beta2_gPP, tau_beta2_gPP)   # prior for effect of disturbance on gammaPP
beta3_gPP[s] ~ dnorm(mu_beta3_gPP, tau_beta3_gPP)  # prior for effect of veg height on gamPP
beta4_gPP[s] ~ dnorm(mu_beta4_gPP, tau_beta4_gPP)   # prior for effect of frate
beta5_gPP[s] ~ dnorm(mu_beta5_gPP, tau_beta5_gPP)   # prior for effect of veg cover
beta6_gPP[s] ~ dnorm(mu_beta6_gPP, tau_beta6_gPP)   # prior for effect of age
}

mu_beta1_gPP ~ dnorm(0, 0.01)
mu_beta2_gPP ~ dnorm(0, 0.01)
mu_beta3_gPP ~ dnorm(0, 0.01)
mu_beta4_gPP ~ dnorm(0, 0.01)
mu_beta5_gPP ~ dnorm(0, 0.01)
mu_beta6_gPP ~ dnorm(0, 0.01)

sigma_beta1_gPP ~ dunif(0, 5)
tau_beta1_gPP <- pow(sigma_beta1_gPP, -2)
sigma_beta2_gPP ~ dunif(0, 5)
tau_beta2_gPP <- pow(sigma_beta2_gPP, -2)
sigma_beta3_gPP ~ dunif(0, 5)
tau_beta3_gPP <- pow(sigma_beta3_gPP, -2)
sigma_beta4_gPP ~ dunif(0, 5)
tau_beta4_gPP <- pow(sigma_beta4_gPP, -2)
sigma_beta5_gPP ~ dunif(0, 5)
tau_beta5_gPP <- pow(sigma_beta5_gPP, -2)
sigma_beta6_gPP ~ dunif(0, 5)
tau_beta6_gPP <- pow(sigma_beta6_gPP, -2)

#detection priors: crew + marker + sex
#prior for effect of crew on p
alpha1_p[1] <- 0 #reference level
for (k in 2:n.crew) {
  alpha1_p[k] ~ dnorm(0, (1/2.25))   # prior on crew effect on logit(p) 
}
#prior for the effect of marker (tag) on p
alpha2_p[1] <- 0 #reference level (No marker)
for (m in 2:3) {
  alpha2_p[m] ~ dnorm(0, (1/2.25))  
}

#prior for effect of sex on p
for(s in 1:n.spp){

beta1_p[s] ~ dnorm(mu_beta1_p, tau_beta_p)   # prior for effect of sex on p

}
mu_beta1_p ~ dnorm(0, 0.01)
sigma_beta_p ~ dunif(0, 5)
tau_beta_p <- pow(sigma_beta_p, -2)


# Priors for intercepts
for (s in 1:n.spp) {
for (t in 1:(n.prim-1)){   
  beta0_phi[t,s] ~ dnorm(mu_beta0_phi, tau_beta0_phi)    #phi intercept
  beta0_gPP[t,s]  ~ dnorm(mu_beta0_gPP, tau_beta0_gPP)   #gPP: prob of staying available given you were available
                                                         #gamPP: probability of emigrating away given you were avail
}
}

mu_beta0_phi ~ dnorm(0, 0.01)
sigma_beta0_phi ~ dunif(0, 5)
tau_beta0_phi <- pow(sigma_beta0_phi, -2)

mu_beta0_gPP ~ dnorm(0, 0.01)
sigma_beta0_gPP ~ dunif(0, 5)
tau_beta0_gPP <- pow(sigma_beta0_gPP, -2)

for (s in 1:n.spp){
for (i in 1:n.ind[s]){
    sex[i,s] ~ dbern(psi)    # imputation
    
    for (t in 1:(n.prim-1)) {
      frate[i,t,s] ~ dnorm(mu_frate, tau_frate)
      disturb[i,t,s] ~ dnorm(mu_disturb, tau_disturb)
      vegheight[i,t,s] ~ dnorm(mu_vegheight, tau_vegheight)
      vegcov[i,t,s] ~ dnorm(mu_vegcov, tau_vegcov)
      age[i,t,s] ~ dbern(psi2)  
    
      logit(phi[i,t,s]) <- beta0_phi[t,s] + beta1_phi[s]*sex[i,s] + beta2_phi[s]*disturb[i,t,s] + beta3_phi[s]*vegheight[i,t,s] + beta4_phi[s]*frate[i,t,s] + beta5_phi[s]*vegcov[i,t,s] + alpha2_phi[mark[i,t,s]]
      logit(gPP[i,t,s]) <- beta0_gPP[t,s] + beta1_gPP[s]*sex[i,s] + beta2_gPP[s]*disturb[i,t,s] + beta3_gPP[s]*vegheight[i,t,s] + beta4_gPP[s]*frate[i,t,s] + beta5_gPP[s]*vegcov[i,t,s] + beta6_gPP[s]*age[i,t,s]
      
    } #t
    
    for (t in 1:(n.prim-2)){
    
      gP[i,t,s] ~ dunif(0,1)
      
    } #t
    } # s
} #i



##########################################################################
# Secondary occasions p's
##########################################################################     
for (s in 1:n.spp){
for (j in 1:3){
    beta0_p[j,s] ~ dnorm(0, (1/2.25))
}
}

for (s in 1:n.spp){
  for (i in 1:n.ind[s]){
    for (t in 1:n.prim) {
      for (j in 1:3){
        logit(p[i,t,j,s]) <- beta0_p[j,s] + alpha1_p[crew[i,t,s]] + alpha2_p[mark[i,t,s]] + beta1_p[s]*sex[i,s]
    }
    }
}
}
  

###########################################################################
# Primary occasions p's
###########################################################################  
for (s in 1:n.spp){
for (i in 1:n.ind[s]) {
  for (t in 2:n.prim){  
    pstar[i,t,s] <- 1 - prod(1 - p[i,t,1:n.sec[t],s])
  }
}
}


# state matrices
for(s in 1:n.spp){
for(i in 1:n.ind[s]){
  sm[i,1,1,1,s] <- phi[i,1,s] * gPP[i,1,s]
  sm[i,1,1,2,s] <- phi[i,1,s] * (1 - gPP[i,1,s])
  sm[i,1,1,3,s] <- 1 - phi[i,1,s]
  sm[i,2,1,1,s] <- 0               
  sm[i,2,1,2,s] <- 1
  sm[i,2,1,3,s] <- 0
  sm[i,3,1,1,s] <- 0
  sm[i,3,1,2,s] <- 0
  sm[i,3,1,3,s] <- 1
}
}

for(s in 1:n.spp){
for(i in 1:n.ind[s]){
    for(t in 2:(n.prim-1)) {
    sm[i,1,t,1,s] <- phi[i,t,s] * gPP[i,t,s]
    sm[i,1,t,2,s] <- phi[i,t,s] * (1 - gPP[i,t,s])
    sm[i,1,t,3,s] <- 1 - phi[i,t,s]
    sm[i,2,t,1,s] <- phi[i,t,s] * gP[i,t-1,s]
    sm[i,2,t,2,s] <- phi[i,t,s] * (1 - gP[i,t-1,s])
    sm[i,2,t,3,s] <- 1 - phi[i,t,s]
    sm[i,3,t,1,s] <- 0
    sm[i,3,t,2,s] <- 0
    sm[i,3,t,3,s] <- 1
    }
}
}

    # observation matrices (o[i, state, t, obs])
for(s in 1:n.spp){
    for (i in 1:n.ind[s]){
      for (t in 2:n.prim){
        o[i,1,t,1,s] <- pstar[i,t,s]
        o[i,1,t,2,s] <- 1 - pstar[i,t,s]
        o[i,2,t,1,s] <- 0
        o[i,2,t,2,s] <- 1
        o[i,3,t,1,s] <- 0
        o[i,3,t,2,s] <- 1
      }
    }
}
  

# Likelihood

 # State process: the latent state
 for(s in 1:n.spp){
 for (i in 1:n.ind[s]){
 
      z[i,first[i,s],s] <- 1  #individuals are state 1 for first cap
      
       for (t in (first[i,s]+1):n.prim){
       # State process: draw S(t) given S(t-1)
         z[i,t,s] ~ dcat(sm[i,z[i,t-1,s],t-1,1:3,s])   
         
      # Observation process: draw O(t) given S(t) using individual-specific o
         ch[i,t,s] ~ dcat(o[i, z[i,t,s], t,1:2,s]) 
        
    
     for (j in 1:n.sec[t-1]){
       mu3[i,t-1,j,s] <- equals(z[i,t-1,s],1) * p[i,t-1,j,s]
       obs[i,t-1,j,s] ~ dbern(mu3[i,t-1,j,s])
  
     } # j
   } # t
 } # i
 } #s
      

########################################################################
#estimating derived parameters
########################################################################

# detection by species
for (s in 1:n.spp){
    mean_pstar[s] <- mean(pstar[1:n.ind[s], 2:n.prim, s])
  }

# detection by month
for (s in 1:n.spp){
  for (j in 1:max(n.sec)){
    mean_p[j,s] <- mean(p[1:n.ind[s], 2:n.prim, j, s])
  }
}

#survival 
#removing 2020 b/c of low sample size (2:(n.prim-1))
#phi by species
for (s in 1:n.spp){
    mean_phi[s] <- mean(phi[1:n.ind[s], 2:(n.prim-1), s])
}

#phi by species & year
for (s in 1:n.spp){
for (t in 2:(n.prim-1)){
    mean_phi_yr[t,s] <- mean(phi[1:n.ind[s], t, s])
}
}

for (s in 1:n.spp) {
  for (m in 1:3) {
    num_tag[s,m]  <- sum(equals(mark[1:n.ind[s], 2:(n.prim-1), s], m))
    sum_phi_tag[s,m] <- sum(phi[1:n.ind[s], 2:(n.prim-1), s] * equals(mark[1:n.ind[s], 2:(n.prim-1), s], m))
    mean_phi_tag[s,m] <- sum_phi_tag[s,m] / num_tag[s,m]
  }
}

#gamma by species
for (s in 1:n.spp){
    mean_gPP[s] <- mean(gPP[1:n.ind[s], 2:(n.prim-1), s])
    mean_gP[s] <- mean(gP[1:n.ind[s], 2:(n.prim-2), s])
    mean_gamPP[s] <- 1 - mean_gPP[s]     # MARK parameterization
    mean_gamP[s] <- 1 - mean_gP[s]     # MARK parameterization
}

#gamma by species & year
for (s in 1:n.spp){
  for (t in 2:(n.prim-1)){
    mean_gPP_yr[t,s] <- mean(gPP[1:n.ind[s], t, s])
    mean_gamPP_yr[t,s] <- 1 - mean_gPP_yr[t,s]     # MARK parameterization
  }
  for (t in 2:(n.prim-2)){
    mean_gP_yr[t,s] <- mean(gP[1:n.ind[s], t, s])
    mean_gamP_yr[t,s] <- 1 - mean_gP_yr[t,s]     # MARK parameterization
  }
}


#Bayes p - checking model fit

 for(s in 1:n.spp){
  for (i in 1:n.ind[s]){
  
      for (t in 1:first[i,s]) {
        ch.new[i,t,s] <- ch[i,t,s]
        }

      for (t in (first[i,s]+1):n.prim){
      
      #Bayes p calc
        ch.new[i,t,s] ~ dcat(o[i, z[i,t,s], t, 1:2, s]) #dmultinom
       }
  }
 }

for(s in 1:n.spp){

  total_new[s] <- sum(ch.new[1:n.ind[s],1:n.prim,s])
  total_obs[s] <- sum(ch[1:n.ind[s],1:n.prim,s])

  bayes.p[s] <- step(total_new[s] - total_obs[s])
}


}", fill = TRUE)
sink()

```

## Model prep and run

```{r prep_run, echo=FALSE}
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#
# MCMC settings
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#

n.chains <- 3
n.thin <- 5
n.adapt <- 10000
n.iter <- 100000
n.burnin <- 15000

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#
# Initial values
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#

#initials by spp.
z.init <- array(NA, dim = dim(ch_array))
start_row <- 1
for (s in 1:n.spp) {
  n_s <- sum(species == s)  # number of individuals for species s
  rows <- start_row:(start_row + n_s - 1)
  
  for (i in 1:n_s) {
    f_i <- f_ann[rows[i]]  # first primary period for this individual
    
    if(f_i < n.prim){
      z.init[i, (f_i + 1):n.prim, s] <- 1
    }
  }
  
  start_row <- start_row + n_s
}

#except for these known states
# BRSP - no known morts, some known temp emig.

z.init[124,3,1] <- 2
z.init[152,3,1] <- 2

#SAGS - no known temp emig, some known mortalities

z.init[124,5,2] <- 3
z.init[58,5,2] <- 3
z.init[108,5,2] <- 3

#SATH

z.init[22,5,3] <- 3
z.init[35,3,3] <- 2


init <- function(){list(z = z.init)} 

z.known <- array(NA, dim = dim(ch_array))

init <- function(){list(z = z.init)} 

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#
# Bundle data
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#
jags_data <- list(n.ind  = c(323, 133, 60),             
                  n.prim = ncol(ch_ann),             
                  n.sec  = n.sec, 
                  ch      = ch_array,        
                  obs    = obs_array,     
                  first  = first_array,        
                  sex = sex_array,
                  crew = crew_array,
                  n.crew = n.crew,
                  disturb = disturb_array,
                  n.spp = n.spp,
                  mark = mark_array,
                  frate = frate_array,
                  vegheight = vegheight_array,
                  vegcov = vegcov_array,
                  age = age_array
)



#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#
# Parameters monitored
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#

pars <- c("mean_pstar", "mean_p","mean_phi", "mean_phi_yr", "mean_phi_tag",
          "mean_gamPP", "mean_gamP", "mean_gamPP_yr", "mean_gamP_yr", 
          "beta1_phi", "beta2_phi", "beta3_phi", "beta4_phi", "beta5_phi", "alpha2_phi", 
          "beta1_gPP", "beta2_gPP", "beta3_gPP", "beta4_gPP", "beta5_gPP", "beta6_gPP",
          "beta1_p", "beta0_phi", "beta0_gPP", "beta0_gP", "bayes.p")

landscape <- jags(data = jags_data, inits = init, parameters.to.save = pars, model.file = "spp_multiveg_age_markov.jags", 
                                          n.chains = n.chains, n.adapt = n.adapt, n.iter = n.iter, n.burnin = n.burnin,
                                          n.thin = n.thin, n.cores = n.chains, parallel = TRUE)

```


## Model assessment


```{r assessment, echo=FALSE}

jagsUI::densityplot(landscape)
jagsUI::traceplot(landscape)

```

