---
title: "Fidelity_territory"
author: "Emily Shertzer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Overview 

This script creates a GLMM used to estimate territory-scale fidelity for the manuscript "Site fidelity of sagebrush songbirds largely invariant with cues of previous nest success or habitat disturbance across spatial scales".

Covariates: disturbance, nest failure rates (plot/year), shrub height, shrub cover

Data: Shertzer, E.R., and Chalfoun, A.D., 2026, Mark recapture and habitat data for adult sagebrush songbirds tracked in Pinedale, Wyoming during 2020 - 2024: U.S. Geological Survey data release, https://doi.org/10.5066/P1EOPE4V.

```{r packages, echo = FALSE}
# clear environment
rm(list = ls())

require(jagsUI)
library(MCMCpack)
library(abind)
library(tidyverse)
library(dplyr)
library(AHMbook)
library(lubridate)
library(ggplot2)
library(bayestestR)

options(scipen = 999)

```

## Data prep

```{r data_prep}
surv <- read.csv("./SageSongbirdSurv_Final.csv", colClasses = c(ch = "character"))

 #fix ages.... so that AHY is NA 
surv <- surv %>%
  dplyr::mutate(age2020 = dplyr::if_else(Age == 1 & Year == 2020, NA, age2020),
      age2021 = dplyr::if_else(Age == 1 & Year == 2021, NA, age2021),
      age2022 = dplyr::if_else(Age == 1 & Year == 2022, NA, age2022),
      age2023 = dplyr::if_else(Age == 1 & Year == 2023, NA, age2023),
      age2024 = dplyr::if_else(Age == 1 & Year == 2024, NA, age2024))

test <- surv %>%
  dplyr::select(BandNumber, Species, Sex, plot,
                mean_utme_2020, mean_utmn_2020,
                mean_utme_2021, mean_utmn_2021, 
                mean_utme_2022, mean_utmn_2022,
                mean_utme_2023, mean_utmn_2023, 
                mean_utme_2024, mean_utmn_2024, 
                kmdistur2020, kmdistur2021, kmdistur2022, kmdistur2023, kmdistur2024,
                frate2020, frate2021, frate2022, frate2023, frate2024,
                at_height_mean2020, at_height_mean2021, at_height_mean2022,
                at_height_mean2023, at_height_mean2024,
                veg_cover_mean2020, veg_cover_mean2021, veg_cover_mean2022, 
                veg_cover_mean2023, veg_cover_mean2024, at_vigor_mean2020,
                at_vigor_mean2021, at_vigor_mean2022, at_vigor_mean2023,
                at_vigor_mean2024,
                age2020, age2021, age2022, age2023,
                depart2020, depart2021, depart2022, depart2023,
                nests2021, nests2022, nests2023
  ) %>%
  dplyr::mutate(
    dist2020_2021 = as.numeric(sqrt((mean_utme_2021 - mean_utme_2020)^2 +
                                      (mean_utmn_2021 - mean_utmn_2020)^2)),
    dist2021_2022 = as.numeric(sqrt((mean_utme_2022 - mean_utme_2021)^2 +
                                      (mean_utmn_2022 - mean_utmn_2021)^2)),
    dist2022_2023 = as.numeric(sqrt((mean_utme_2023 - mean_utme_2022)^2 +
                                      (mean_utmn_2023 - mean_utmn_2022)^2)),
    dist2023_2024 = as.numeric(sqrt((mean_utme_2024 - mean_utme_2023)^2 +
                                      (mean_utmn_2024 - mean_utmn_2023)^2))
  )

plotdat <- test %>%
  transmute(
    BandNumber,
    Species,
    Sex,
    plot,
    year = 2020, 
    kmdistur = kmdistur2020,
    frate = frate2020,
    distance = dist2020_2021,
    vegheight = at_height_mean2020,
    vegcov = veg_cover_mean2020,
    vegvig = at_vigor_mean2020,
    age = age2020,
    depart = depart2020,
    nest = NA
  ) %>%
  bind_rows(
    test %>%
      transmute(
        BandNumber,
        Species,
        Sex,
        plot,
        year = 2021,
        kmdistur = kmdistur2021,
        frate = frate2021,
        distance = dist2021_2022,
        vegheight = at_height_mean2021,
        vegcov = veg_cover_mean2021,
        vegvig = at_vigor_mean2021,
        age = age2021,
        depart = depart2021,
        nest = nests2021
      ),
    test %>%
      transmute(
        BandNumber,
        Species,
        Sex,
        plot,
        year = 2022,
        kmdistur = kmdistur2022,
        frate = frate2022,
        distance = dist2022_2023,
        vegheight = at_height_mean2022,
        vegcov = veg_cover_mean2022,
        vegvig = at_vigor_mean2022,
        age = age2022,
        depart = depart2022,
        nest = nests2022
      ),
    test %>%
      transmute(
        BandNumber,
        Species,
        Sex,
        plot,
        year = 2023,
        kmdistur = kmdistur2023,
        frate = frate2023,
        distance = dist2023_2024,
        vegheight = at_height_mean2023,
        vegcov = veg_cover_mean2023,
        vegvig = at_vigor_mean2023,
        age = age2023,
        depart = depart2023,
        nest = nests2023
      )
  )

#prep data for jags

mod_dat <- plotdat %>%
  dplyr::filter(!is.na(distance)) %>%
  dplyr::mutate(sex = case_when(Sex == "M" ~ 1, Sex == "F" ~ 0, Sex == "U" ~ NA_real_))

mod_dat <- mod_dat %>%
  mutate(
    species_id = as.integer(as.factor(Species)),
    band_id = as.integer(as.factor(BandNumber)),
    year_id = as.integer(as.factor(year)),
    
    # Covariates: center/scaled
    kmdistur_z = scale(kmdistur)[,1],
    frate_z = scale(frate)[,1],
    vegheight_z = scale(vegheight)[,1],
    vegcov_z = scale(vegcov)[,1],
    vegvig_z = scale(vegvig)[,1],
    bin_age = dplyr::case_when(is.na(age) ~ NA, age == 2 ~ 1, TRUE ~ 0)) #SY = 1, ASY = 0

mod_dat$logdist <- log(mod_dat$distance)
mod_dat$logdist[127] <- 0  #problem value

# changing SATH ages to NA
mod_dat <- mod_dat %>%
  dplyr::mutate(bin_age = ifelse(Species == "SATH", NA, bin_age))

```

## Model

```{r glmm_model, echo=FALSE}
sink("site_fidelity_multiveg_yrint.jags")
cat("
model {

  ############ Priors ############
  
  # data imputation
  psi ~ dunif(0,1)  #sex ratio
  psi2 ~ dunif(0,1)  #age ratio
  
  mu_frate ~ dnorm(0, 0.01)   
  tau_frate <- pow(sd_frate, -2)
  sd_frate ~ dunif(0, 5)
  
  mu_vegheight ~ dnorm(0, 0.01) 
  tau_vegheight <- pow(sd_vegheight, -2)
  sd_vegheight ~ dunif(0, 5)
  
  mu_vegcov ~ dnorm(0, 0.01) 
  tau_vegcov <- pow(sd_vegcov, -2)
  sd_vegcov ~ dunif(0, 5)

  # Species-level slopes
  mu_beta1 ~ dnorm(0, 0.01)
  mu_beta2 ~ dnorm(0, 0.01)
  mu_beta3 ~ dnorm(0, 0.01)
  mu_beta4 ~ dnorm(0, 0.01)
  mu_beta5 ~ dnorm(0, 0.01)
  mu_beta6 ~ dnorm(0, 0.01)

  sigma_beta1 ~ dunif(0, 5)
  tau_beta1 <- pow(sigma_beta1, -2)
  sigma_beta2 ~ dunif(0, 5)
  tau_beta2 <- pow(sigma_beta2, -2)
  sigma_beta3 ~ dunif(0, 5)
  tau_beta3 <- pow(sigma_beta3, -2)
  sigma_beta4 ~ dunif(0, 5)
  tau_beta4 <- pow(sigma_beta4, -2)
  sigma_beta5 ~ dunif(0, 5)
  tau_beta5 <- pow(sigma_beta5, -2)
  sigma_beta6 ~ dunif(0, 5)
  tau_beta6 <- pow(sigma_beta6, -2)

  for(s in 1:n_spp){
    beta1[s] ~ dnorm(mu_beta1, tau_beta1)   # sex effect by species
    beta2[s] ~ dnorm(mu_beta2, tau_beta2)   # frate effect by species
    beta3[s] ~ dnorm(mu_beta3, tau_beta3)   # disturbance effect by species
    beta4[s] ~ dnorm(mu_beta4, tau_beta4)   # veg height effect by species
    beta5[s] ~ dnorm(mu_beta5, tau_beta5)   # veg cover effect by species
    beta6[s] ~ dnorm(mu_beta6, tau_beta6)   # age effect by species
  }


  # Species-level intercepts
  mu_beta0 ~ dnorm(0, 0.01)
  sigma_beta0 ~ dunif(0, 5)
  tau_beta0 <- pow(sigma_beta0, -2)

for (t in 1:n_prim){
  for(s in 1:n_spp){
    beta0[t,s] ~ dnorm(mu_beta0, tau_beta0)
  }
}

  # Random Effect of individual
  sigma_ind ~ dunif(0, 5)
  tau_ind <- pow(sigma_ind, -2)
  
  for(i in 1:n_ind){
    ind_re[i] ~ dnorm(0, tau_ind)
  }

  #sex imputation
  
  for (i in 1:n_ind){
    sex[i] ~ dbern(psi)
}


  ############ Likelihood ############

  for(n in 1:n_obs){
  #imputing missing data
    age[n] ~ dbern(psi2)
    frate_z[n] ~ dnorm(mu_frate, tau_frate) 
    vegheight_z[n] ~ dnorm(mu_vegheight, tau_vegheight)
    vegcov_z[n] ~ dnorm(mu_vegcov, tau_vegcov)
  
    distance[n] ~ dnorm(mu[n], tau)    # response model
    
    mu[n] <- beta0[year_id[n],species_id[n]] +
             beta1[species_id[n]] * sex[band_id[n]] +
             beta2[species_id[n]] * frate_z[n] +
             beta3[species_id[n]] * kmdistur_z[n] +
             beta4[species_id[n]] * vegheight_z[n] +
             beta5[species_id[n]] * vegcov_z[n] +
             beta6[species_id[n]] * age[n] +
             ind_re[band_id[n]]
  }

  sigma ~ dunif(0, 5)
  tau <- pow(sigma, -2)
  
#Bayesian p-value
  #copying kery royle vol 1, pg 193:
  #chi square: sum of (obs-exp) ^ 2 / exp

# Fit assessments: Chi-square test statistic and posterior predictive check
for(n in 1:n_obs){

chi2[n] <- pow((distance[n]-mu[n]),2) / (sqrt(mu[n])+0.0001) # obs.
dist.new[n] ~ dnorm(mu[n], tau) # Replicate (new) data set
chi2.new[n] <- pow((dist.new[n]-mu[n]),2) / (sqrt(mu[n])+0.0001) # exp.
}
# Add up discrepancy measures for entire data set
fit <- sum(chi2[]) # Omnibus test statistic actual data
fit.new <- sum(chi2.new[]) # Omnibus test statistic replicate data


}", fill = TRUE)
sink()

```

## Model prep and run

```{r prep_run, echo=FALSE}
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#
# MCMC settings
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#

n.chains <- 3
n.thin <- 5
n.adapt <- 10000
n.iter <- 100000
n.burnin <- 15000

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#
# Bundle data
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#

data = list(
  n_obs = nrow(mod_dat),
  n_ind = length(unique(mod_dat$BandNumber)),
  n_prim = length(unique(mod_dat$year_id)),
  n_spp = length(unique(mod_dat$species_id)),
  distance = mod_dat$logdist,
  frate_z = mod_dat$frate_z,
  kmdistur_z = mod_dat$kmdistur_z,
  vegheight_z = mod_dat$vegheight_z,
  vegcov_z = mod_dat$vegcov_z,
  species_id = mod_dat$species_id,
  band_id = mod_dat$band_id,
  year_id = mod_dat$year_id,
  sex = mod_dat$sex,
  age = mod_dat$bin_age
)

#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#
# Parameters monitored
#%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%#

params <- c("beta0","beta1","beta2","beta3", "beta4", "beta5", "beta6",
            "sigma_ind","sigma_year","sigma_beta","sigma_beta0","psi",
            "dist.new", "T_obs", "T_rep", "fit", "fit.new")


territory_fidel1 <- jags(data = data, parameters.to.save = params, model.file = "site_fidelity_multiveg_yrint.jags", 
                 n.chains = n.chains, n.adapt = n.adapt, n.iter = n.iter, n.burnin = n.burnin,
                 n.thin = n.thin, n.cores = n.chains,  parallel = TRUE)

```


## Model assessment


```{r assessment, echo=FALSE}

jagsUI::densityplot(territory_fidel1)
jagsUI::traceplot(territory_fidel1)

#Check model fit

# Check for overfitting (make sure there's some variation btwn real and sim values)

dist_new_mean <- apply(territory_fidel1$sims.list$dist.new, 2, mean)

plot(mod_dat$logdist, dist_new_mean,
     xlab = "Observed distance",
     ylab = "Posterior predictive mean",
     pch = 16, col = rgb(0, 0, 0, 0.5))
abline(0, 1, lwd = 2)

i <- sample(seq_len(nrow(territory_fidel1$sims.list$dist.new)), 1)
plot(mod_dat$logdist,
     territory_fidel1$sims.list$dist.new[i, ],
     xlab = "Observed",
     ylab = "Replicated")
abline(0,1)

# check bayesian p-values for under-fitting

bpv <- mean(territory_fidel1$sims.list$fit.new > territory_fidel1$sims.list$fit)
bpv

```

